---
title: "NARFD Example"
author: "Bohan Wu"
date: "3/15/2021"
output: pdf_document
---
```{r}
rm(list = ls())
```
```{r setup, include=FALSE}
source(here::here("source", "Helper Functions.R"))
source(here::here("source", "NARFD.R"))

library(tidyverse)
library(reshape2)
library(WeightedROC)
knitr::opts_chunk$set(
	echo = TRUE,
  cache=FALSE,
	warning = FALSE,
	fig.width = 8, 
  fig.height = 6,
  out.width = "90%"
)

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d

theme_set(theme_minimal() + theme(legend.position = "bottom"))
```
This document includes a brief demonstration of using Guassian FPCA, NARFD, Poisson FPCA to extract features from the preprocessed "Nhanes" activity count data aggregated in 5-min intervals. The data was generated using methods included in the file "process_nhanes_data.R". 

## Example data

Observations are included in `./data/pa_nhanes_5min.RData`. These data are imported and plotted in the code chunk below.

```{r}
load(here::here("data", "pa_nhanes_5min2.RData"))

rownames(Y) <- seq(nrow(Y)); colnames(Y) <- seq(ncol(Y))
```
Plot the activity curves
```{r}
{
par(mfrow = c(1,2))
apply(Y,1,mean) %>% log() %>% hist(ylab = "log(E(y_i))")

apply(Y,1,var) %>% log() %>% hist(ylab = "log(var(y_i))")
}

melt(Y) %>% 
  dplyr::rename(.id = Var1, .index = Var2, .observed = value)%>%
  mutate(.id = as.factor(.id)) %>% 
  ggplot(aes(x = .index, y = .observed, group = .id)) +
  geom_line()


```
## Fit NARFD and PFPCA

Next we fit the proposed method using the `NARFD()` function and a pre-specified penalty sequence. In other settings, a wider range of potential penalty terms and larger number of CV folds may be more appropriate.

###NARFD
```{r}
Y.l<- MakeLong(t(Y))
penalty.seq <- 2 ^ c(-3, 1, 3,5)
#nbas = max(20,min(ceiling(ncol(Y)/4), 150))

narfd_results <- NARFD(long = Y.l, npc = 6, nbasis = 25, D = ncol(Y), type = "NARFD", periodic = FALSE, folds = 3, iter = 30,  penalty.seq = penalty.seq, verbose = TRUE)
```
###PFPCA
```{r}
Y.l<- MakeLong2(Y)
 # pfpca_results <- NARFD(long = Y.l, npc = 6, nbasis = 25, D = ncol(Y), type = "GFPCA", periodic = FALSE, folds = 3, iter = 20,  penalty.seq = penalty.seq, verbose = TRUE)

pfpca_results <- gfpca_twoStep(Y.l, family = "poisson", npc = 6,Kt = 10)


rm(list = c('Y_m', 'Y.l'))
```

```{r}
print(paste("optimal penalty = ",narfd_results$penalty))
```
The plot below shows estimated functional prototypes.
```{r}
fpca_fit <- fpca.face(log(Y + 1),npc = 6)

F_gaus <- MakeLong(exp(fpca_fit$efunctions[,1:6])-1) %>%
  mutate(type = "FPCA") %>%
  rename(prototype = .id, Value = .observed)

F_pois <- MakeLong(exp(pfpca_fit$efunctions[,1:6])-1) %>%
  mutate(type = "PFPCA") %>%
  rename(prototype = .id, Value = .observed)

F_narfd <- narfd_results$prototypes %>% mutate(type = "NARFD")

F_tot <- rbind(rbind(F_gaus[,c("prototype",".index","Value","type")],
               F_pois[,c("prototype",".index","Value","type")]),
               F_narfd)

F_tot$prototype <- gsub(c("IV","prototype"),"FPC ",F_tot$prototype)

F_tot$prototype <- gsub("prototype","FPC ",F_tot$prototype)

{
#jpeg(file.path("../plots","functional_prototypes.jpeg"),height=900,width=1350,quality = 100)
 p =  ggplot(data = F_tot, aes(x = .index, y = Value)) + 
  geom_line() + 
  facet_grid(type~prototype) + 
  scale_x_continuous(breaks = c(2,8,14,20)*12 +1, labels = c("02:00","08:00","14:00","20:00"))+
  theme(axis.text.x = element_text(angle=45),text = element_text(size = 35))+
  xlab("Time")
print(p)
rm(list = c("p", "F_gaus","F_pois","F_narfd","F_tot"))
#dev.off()
}
```

Lastly we show the observed data and fitted value for four activity curves. 

1)Extract Yhat for Gaussian FPCA and PFPCA. 
```{r,warning=FALSE}
#Yhat: Gassian FPCA
fpca_pred <- MakeLong(t(exp(fpca_fit$Yhat)-1)) %>% rename(val.fpca = .observed)

#Yhat: Poisson FPCA
f.pred <- function(y){
   fit.y <-  glm.fit(as.matrix(pfpca_fit$efunctions), y, family = poisson())
   fit.y$fitted.values
   
}
pfpca_pred <- MakeLong(apply(Y,1,f.pred)) %>% rename(val.pfpca = .observed)
```
2)Plot the observed data and fitted values for four activity curves. 
```{r}
n_samples = 6
samples = sample(unique(narfd_results$pred$.id),n_samples)

p= vector("list",n_samples)
samples = paste0("I",c(700,680, 888, 930, 305, 635))

{
jpeg(file.path("../plots","Example Fit.jpeg"),height=900,width=1350,quality = 100)
for(i in seq(n_samples)){
  
  dd =   narfd_results$pred[,c(".index",".id",".observed",".pred")] %>%       inner_join(fpca_pred,by = c(".index",".id")) %>% 
    inner_join(pfpca_pred,by = c(".index",".id")) %>% 
    filter(.id == samples[i])
  
  p[[i]] =ggplot(data = dd, aes(x = .index, y = .pred)) + 
    geom_line(aes(color = "NARFD"),size = 1.5) + 
    geom_line(aes(y = val.pfpca,color = "Poisson FPCA"),size = 1.5) + 
    geom_line(aes(y = val.fpca,color = "Gaussian FPCA"),size = 1.5) + 
    geom_point(aes(y = .observed)) +
    ggtitle(str_replace(samples[i],"IV","Subject "))+
    scale_color_manual("", values=c("NARFD" = "blue", "Gaussian FPCA" = "gold","Poisson FPCA" = "red"))+ 
    scale_x_continuous(breaks = c(2,8,14,20)*12 +1, labels = c("02:00","08:00","14:00","20:00"))+
    theme(axis.text.x = element_text(angle=45),text = element_text(size = 25))+
    xlab("Time")
  
}
  
  multiplot(p[[1]],p[[2]],p[[3]],p[[4]],p[[5]], p[[6]],  cols=3)
dev.off()
}


#rm(list = c("p", "dd","fpca_pred","pfpca_pred","f.pred"))
```

```{r}
# Save the results:
save(narfd_results, file = 'NARFD_results.RData')
save(pfpca_results, file = 'PFPCA_results.RData')
save(fpca_fit, file = 'GFPCA_results.RData')

```

##Create Table 1
```{r}
source(here::here("source", "create_table1.R"))

rm(list = c("scores_m","samples","roc_ROCR","AUC"))
```
