---
title: "NARFD Example"
author: "Bohan Wu"
date: "3/15/2021"
output: pdf_document
---
```{r}
rm(list = ls())
```
```{r setup, include=FALSE}
source(here::here("source", "Helper Functions.R"))
source(here::here("source", "NARFD.R"))

library(tidyverse)
library(reshape2)
library(WeightedROC)
library(ggpubr)
knitr::opts_chunk$set(
	echo = TRUE,
  cache=FALSE,
	warning = FALSE,
	fig.width = 8, 
  fig.height = 6,
  out.width = "90%"
)

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d

theme_set(theme_minimal() + theme(legend.position = "right"))
```
This document includes a brief demonstration of using Guassian FPCA, NARFD, Poisson FPCA to extract features from the preprocessed "Nhanes" activity count data aggregated in 5-min intervals. The data was generated using methods included in the file "process_nhanes_data.R". 

## Example data

Observations are included in `./data/pa_nhanes_5min.RData`. These data are imported and plotted in the code chunk below.

```{r}
load(here::here("data", "pa_nhanes_5min2.RData"))

rownames(Y) <- seq(nrow(Y)); colnames(Y) <- seq(ncol(Y))
```
Plot the activity curves
```{r}
{
par(mfrow = c(1,2))
apply(Y,1,mean) %>% log() %>% hist(ylab = "log(E(y_i))")

apply(Y,1,var) %>% log() %>% hist(ylab = "log(var(y_i))")
}

melt(Y) %>% 
  dplyr::rename(.id = Var1, .index = Var2, .observed = value)%>%
  mutate(.id = as.factor(.id)) %>% 
  ggplot(aes(x = .index, y = .observed, group = .id)) +
  geom_line()


```
## Fit NARFD and PFPCA

Next we fit the proposed method using the `NARFD()` function and a pre-specified penalty sequence. In other settings, a wider range of potential penalty terms and larger number of CV folds may be more appropriate.

###NARFD
```{r}
Y.l<- MakeLong(t(Y))
penalty.seq <- 2 ^ c(-3, 1, 3,5)
#nbas = max(20,min(ceiling(ncol(Y)/4), 150))

narfd_results <- NARFD(long = Y.l, npc = 6, nbasis = 25, D = ncol(Y), type = "NARFD", periodic = FALSE, folds = 3, iter = 30,  penalty.seq = penalty.seq, verbose = TRUE)
```
###PFPCA
```{r}
Y.l<- MakeLong2(Y)

pfpca_results <- gfpca_twoStep(Y.l, family = "poisson", npc = 6,Kt = 10)

rm(list = c('Y_m', 'Y.l'))
```

```{r}
print(paste("optimal penalty = ",narfd_results$penalty))
```
The plot below shows estimated functional prototypes.
```{r}
fpca_fit <- fpca.face(log(Y + 1),npc = 6)

F_gaus <- MakeLong(exp(fpca_fit$efunctions[,1:6])-1) %>%
  mutate(type = "GFPCA") %>%
  dplyr::rename(prototype = .id, Value = .observed)

F_pois <- MakeLong(exp(pfpca_results$efunctions[,1:6])-1) %>%
  mutate(type = "PFPCA") %>%
  dplyr::rename(prototype = .id, Value = .observed)

F_narfd <- narfd_results$prototypes %>% mutate(type = "NARFD")

F_tot <- rbind(rbind(F_gaus[,c("prototype",".index","Value","type")],
               F_pois[,c("prototype",".index","Value","type")]),
               F_narfd)

F_tot$prototype <- gsub(c("IV","prototype"),"FPC ",F_tot$prototype)

F_tot$prototype <- gsub("prototype","FPC ",F_tot$prototype)

{
jpeg(file.path("../plots","functional_prototypes.jpeg"),height=900,width=1350,quality = 100)
 p =  ggplot(data = F_tot, aes(x = .index, y = Value)) + 
  geom_line(size = 3) + 
  facet_grid(type~prototype) + 
  scale_x_continuous(breaks = c(2,8,14,20)*12 +1, labels = c("02:00","08:00","14:00","20:00"))+
  theme(axis.text.x = element_text(angle=45),text = element_text(size = 40))+
  xlab("Time")
print(p)
rm(list = c("p", "F_gaus","F_pois","F_narfd","F_tot"))
dev.off()
}
```

Lastly we show the observed data and fitted value for four activity curves. 

1)Extract Yhat for Gaussian FPCA and PFPCA. 
```{r,warning=FALSE}
#Yhat: Gassian FPCA
{
fpca_pred <- t(exp(fpca_fit$Yhat)-1);colnames(fpca_pred) <- 1:nrow(Y)
fpca_pred <- MakeLong(fpca_pred) %>% rename(val.fpca = .observed)
}

#Yhat: Poisson FPCA
{
f.pred <- function(y)  glm.fit(as.matrix(pfpca_results$efunctions), y, family = poisson())$fitted.values
pfpca_pred <- apply(Y,1,f.pred);colnames(pfpca_pred) <- 1:nrow(Y)
pfpca_pred <- MakeLong(pfpca_pred) %>% rename(val.pfpca = .observed)
}

#Yhat:NARFD
narfd_pred <- narfd_results$pred[,c(".index",".id",".observed",".pred")]
```
2)Plot the observed data and fitted values for four activity curves. 
```{r}
n_samples = 4

p= vector("list",n_samples)
#samples = paste0("I",c(906,680, 888, 708, 305, 1))
samples = paste0("I",c(680, 888, 708,1))

#samples = paste0("I",sample(seq(nrow(X)),6))
{
jpeg(file.path("../plots","Example Fit poster.jpeg"),height=800,width=1100,quality = 100)
for(i in seq(n_samples)){
  dd =   narfd_pred%>%       
    inner_join(fpca_pred,by = c(".index",".id")) %>% 
    inner_join(pfpca_pred,by = c(".index",".id")) %>% 
    filter(.id == samples[i])
  
  p[[i]] =ggplot(data = dd, aes(x = .index, y = .pred)) + 
    geom_line(aes(color = "NARFD"),size = 2.5) +
    geom_line(aes(y = val.fpca,color = "Gaussian FPCA"),size = 2.5) + 
    geom_line(aes(y = val.pfpca,color = "Poisson FPCA"),size = 2.5) + 
    geom_point(aes(y = .observed),size = 2.5) +
 #   ggtitle(str_replace(samples[i],"I","Subject "))+
    scale_color_manual("", values=c("NARFD" = "blue", "Gaussian FPCA" = "green","Poisson FPCA" = "red"))+ 
    scale_x_continuous(breaks = c(2,8,14,20)*12 +1, labels = c("02:00","08:00","14:00","20:00"))+
    theme(axis.text.x = element_text(angle=45),text = element_text(size = 35))+
    labs(x = "Time", y = "Y")
}
  pp <- ggarrange(p[[1]],p[[2]],p[[3]],p[[4]], nrow=2,ncol = 2, common.legend = TRUE, legend="bottom", heights = c(3,3.75))
print(pp)
dev.off()
}

#rm(list = c("p", "dd","fpca_pred","pfpca_pred","f.pred","pp"))
```

```{r}
# Save the results:
save(narfd_results, file = 'NARFD_results.RData')
save(pfpca_results, file = 'PFPCA_results.RData')
save(fpca_fit, file = 'GFPCA_results.RData')

```

#Residuals
```{r}
r1 <- rowMeans(abs(fpca_fit$Yhat - Y)) 
r2 <- rowMeans(abs(t(apply(Y,1,f.pred)) - Y))  
r3 <- rowMeans(abs(reshape2::dcast(narfd_pred, .id ~ .index,value.var = ".pred") %>% column_to_rownames(".id")  - Y)) 

res_df <- data.frame(residual = c(r1,r2,r3), type = rep(c("GFPCA","PFPCA","NARFD"), each = nrow(Y))) 
getwd()
jpeg(file.path("../plots","FPCA MAE Plot.jpeg"),height=800,width=1350,quality = 100)
ggplot(res_df, aes(x = fct_reorder(type, residual, .fun='mean'), y = residual)) +   
    coord_flip() + # Applying ggplot function
  geom_boxplot()+
  labs(y = "MAE", x = "Method")+
  ggtitle("Mean Absolute Errors (MAE) for GFPCA, NARFD, PFPCA")+
    theme(text = element_text(size = 27))
dev.off()
rm(list = c("r1","r2","r3","res_df"))
```

