---
title: "NARFD example"
subtitle: "Using simulated accelerometer data"
output: html_document
---

```{r setup, include=FALSE}
source(here::here("source", "NARFD.R"))

library(tidyverse)

knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
	fig.width = 8, 
  fig.height = 6,
  out.width = "90%"
)

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d

theme_set(theme_minimal() + theme(legend.position = "bottom"))
```

This document includes a brief illustration of the NARFD method for ...

## Example data

Observations are included in `./data/accel_df.csv`. These data are imported and plotted in the code chunk below.

```{r}
accel_df = 
  read_csv(here::here("data", "accel_df.csv"))

accel_df %>% 
  ggplot(aes(x = .index, y = .observed, group = .id)) +
  geom_line(alpha = .2)
```


## Fit NARFD

Next we fit the proposed method using the `NARFD()` function and a pre-specified penalty sequence. In other settings, a wider range of potential penalty terms and larger number of CV folds may be more appropriate.

```{r, fig.height = 4}
penalty.seq <- 2 ^ c(-5, -3, -1, 1, 3, 5)

narfd_results <- 
  NARFD(long = accel_df, npc = 5, D = 144, type = "NARFD", nbasis = 35,
        periodic = TRUE, folds = 3, iter = 25, penalty.seq = penalty.seq)
```

The plot below shows estimated functional prototypes.

```{r}
narfd_results$prototypes %>% 
  ggplot(aes(x = .index, y = Value)) + 
  geom_line() + 
  facet_grid(~prototype) + 
  xlab("Time index")
```

Lastly we show the observed data and fitted value for one subject.

```{r}
narfd_results$pred %>% 
  filter(.id == "subj_1") %>% 
  ggplot(aes(x = .index, y = .pred)) + 
  geom_line() + 
  geom_point(aes(y = .observed)) +
  xlab("Time index")
```

