---
title: "NARFD example"
subtitle: "Using artificial data"
output: html_document
---

```{r}
#install.packages(c("nloptr", "splines", "pbs", "refund", "NNLM", "dplyr", "tidyr", "data.table", "glmnet", "mgcv", "tibble"))

install.packages("NNLM")
library(devtools)
install_github('linxihui/NNLM')
```
```{r setup, include=FALSE}
source(here::here("source", "NARFD.R"))

library(tidyverse)

knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
	fig.width = 8, 
  fig.height = 6,
  out.width = "90%"
)

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d

theme_set(theme_minimal() + theme(legend.position = "bottom"))
```

This document includes a brief illustration of the NARFD method for "artificial" simulated data, generated using sine and cosine functional prototypes in a way that is similar to the full simulation exercise in the manuscript. 

## Example data

Observations are included in `./data/artificial_example.csv`. These data are imported and plotted in the code chunk below.

```{r}
artificial_df = 
  read_csv(here::here("data", "artificial_example.csv"))

artificial_df %>% 
  ggplot(aes(x = .index, y = .observed, group = .id)) +
  geom_line()
```

These were generated using the code in the next chunk and exported to a csv file. The `Simulate()` function produces a list that includes observed data as well as true prototypes, scores, and latent means. The `MakeLong()` function extracts only observed data and converts this to a long-form dataframe suitable for fitting. 

```{r, eval = FALSE}
source(here::here("source", "Simulate_Data.R"))
set.seed(1)

artificial_data <- Simulate(N = 20, D = 80, type = "NARFD", seed = 1)

dim(artificial_data$observations)
artificial_df <- MakeLong(artificial_data$observations)
```

## Fit NARFD

Next we fit the proposed method using the `NARFD()` function and a pre-specified penalty sequence. In other settings, a wider range of potential penalty terms and larger number of CV folds may be more appropriate.

```{r}
penalty.seq <- c(0.1, 1, 10)

narfd_results <- 
  NARFD(long = artificial_df, npc = 2, D = 80, type = "NARFD", nbasis = 20,
        periodic = FALSE, folds = 3, iter = 25, penalty.seq = penalty.seq)
```

The plot below shows estimated functional prototypes.

```{r}
narfd_results$prototypes %>% 
  ggplot(aes(x = .index, y = Value, col = prototype)) + 
  geom_line() + 
  xlab("Time index")
```

Finally we show the observed data and fitted value for one subject.

```{r}
narfd_results$pred %>% 
  filter(.id == "IV1") %>% 
  ggplot(aes(x = .index, y = .pred)) + 
  geom_line() + 
  geom_point(aes(y = .observed)) +
  xlab("Time index")
```

